---
title: "Data import"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Read in some data

Absolute path: start from very beginning to desired folder (do not use this method) Start from harddrive-> user-> nidhipatel-> dekstop-> datasciencei-> datawanglingi->data->file. This would only apply to my computer. 
Relative path: start where you are and go where you need to go, which is prob close by. (use this only) Start at data_import.Rmd -> data folder -> file. Can be reproduced in anyones computer

Readn in the litters dataset.

```{r}
litters_df <- read_csv("./data/FAS_litters.csv")
# the "." says starting in this folder.
litters_df = janitor::clean_names(litters_df)
#clean_names fn is not in tidyverse, but in janitors package. To load this fn without loading the entire package we use "::" to pull it from the package. 
```


## Take a look at data
Printing in console.
```{r}
litters_df
```

```{r}
head(litters_df)
#gives first 6 rows. 
tail(litters_df)
#gives the last few rows.

skimr::skim(litters_df)
#tells you how many are missing, what are the white spaces, mean, std, min, max, histograms. 

#view(litters_df) to open. but only use in console. (which we already knew)

```

## Options to read_csv

check out `?read_csv` for more info
```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE) #col_names= FALSE makes up col names

#litters_df = read_csv("./data/FAS_litters.csv", na = c("", "NA", ".", 999)) #replaces all those with NA

litters_df = read_csv("./data/FAS_litters.csv", )

```

look at read_csv


## Other file formats
read in excel file.

```{r}
library(readxl) #next time keep all loaded packages at the top of the code.

mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```

Read in a SAS file

```{r}
library(haven)

pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```


## Comparison with Base R.
what about `read.csv`..?
base R is more difficult to look at 

```{r}
litters_base = read.csv("data/FAS_litters.csv")
litters_readr = read_csv("data/FAS_litters.csv")

litters_base
litters_readr

```

