---
title: "Visualizations"
subtitle: "Part 1"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors( 
    c("USW00094728", "USC00519397", "USS0023B17S"), 
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
#rnoaa package allows you to pull public data from online. 
#want data drom these three weather stations
    mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  #the degrees that come from rnoaa are in a 10ths of a deg C. 
  select(name, id, everything())

weather_df

```


## Scatterplots!!

Create my first scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

New approach, same plot

```{r}
weather_df %>% #allows to add mutate, select, etc
ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Save and edit a plot object

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```


## Advanced scatterplot

Start with same one and make it fancy! 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE) #adds nice smooth curves that run through data set (i do want to know how they are calculated tho)
```

What about the `aes` placement 

```{r}
 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) +
  geom_smooth(se = FALSE) #we only see one blue line, not separated by colors.  this color only applies to the scatterplot, bc the asthetic is added to point. define asthetics in ggplot and they get passed on to everything below.

```

Let's facet some things!!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + #if i add alpha = tmin in the aes here, it would gradient for the min levels to be more transparent
  geom_point(alpha = .02, size = 3) + #alpha is transparency, size is size of points; can be used in geom_smooth too. 
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name) # . says don't create (rows in this case), ~ separates (x ~ y), y creates different columns according to names. 

```

Let's combine some elements and try a new plot.

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = 0.5) + 
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)
#make a v simple scatterplot first and tweak as time goes on. 
```

## Small notes

How many geoms have to exist?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_smooth(se = FALSE)
```

do not need the points, I can have any geoms I want

I can use neat geom

```{r}
 weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  #geom_hex() #geom_hex won't work for me.
  #geom_bin2d() same as hex, but with squares.
  geom_density2d() +
  geom_point(alpha = 0.3)
```

## Univariate plots

Histograms are really great. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) + 
  geom_histogram()
```

Can we add color

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + #used fill instead of color
  geom_histogram(position = "dodge") + #to dodge instead of stack the histograms.
  facet_grid(.~name)
```

Let's try a new geometry

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_density(alpha = 0.3, adjust = 0.5)
```

What about box plots?

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) + 
  geom_boxplot()
```

Trendy plots

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_violin(alpha = 0.4) + #cross between boxplot and density plot.
  stat_summary()
```

Ridge plot- most popular plot od 2017

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) + 
  geom_density_ridges() #density curves. load from ggridges package. 
```

